#Genotype processing WMB
```{r}
library(readr)
library(readxl)
library(stringr)
library(dplyr)
```

```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "/home/karl/git/BarleyGermination-KK")  # with something else than `getwd()`
getwd()
```
#add AlaAT for DH population
AlaT_markers<-read_excel("data/Genotype_data/WMB_DH_AlaAT_KASP_rawdata.xlsx",sheet = 'Results');AlaT_markers[,c(1:12)];colnames(AlaT_markers)[12]<-"AlaT_Allele"
AlaT_markers<-AlaT_markers[!is.na(AlaT_markers$AlaT_Allele),];colnames(AlaT_markers)[4]<-"GID"

#remove het called markers, most likely a mistake
#AlaT_markers$AlaT_Allele<-as.factor(AlaT_markers$AlaT_Allele)
#cant be a factor once its in the genotype table

AlaT_markers<-AlaT_markers%>%as.data.frame()%>%select(GID,Allele_bp)
AlaT_markers[AlaT_markers$Allele_bp%in%c("G/C"),]<-AlaT_markers[AlaT_markers$Allele_bp%in%c("G/C"),]
AlaT_markers<-AlaT_markers%>%filter(Allele_bp%in%c("G","C"))%>%as.data.frame();row.names(AlaT_markers)<-AlaT_markers$GID
AlaT_markers_1<-as.data.frame(t(as.matrix(AlaT_markers)));AlaT_markers_1<-AlaT_markers_1[-1,]
AlaT_markers_1<-AlaT_markers_1%>%as.data.frame%>%tibble::add_column("rs#"="Qsd1","alleles"="G/C","chrom"="5H","pos"=442160000,"strand"="+",.before =1)
rownames(AlaT_markers_1)<-NULL
```
#input the new data
```{r}
wmb_gt<- read_delim("data/Genotype_data/all_filter.hmp.txt", 
                 delim = "\t", escape_double = FALSE, 
                trim_ws = TRUE)
str(wmb_gt)

### Processing

#remove DH130910 and SY tepee duplicates since they were in the RIL pop, also remove lines not relevants
#for this population like Buck, 1-6 or DH130935
colnames(wmb_gt)
#d_names<-c("BS710-65","BS713-79","BS713-103","BS715-132","BS911-50","BS911-58")
#this was figured out after removing the number prefix and finding the duplicate names, we want
#the genotypes in the RIL as they are the most recent
#colnames(wmb_gt[grep(d_names[6], names(wmb_gt), value = TRUE)])
#exclude_names
#exclude_names<-c("186_BS710-65","187_BS713-79","188_BS713-103","211_BS715-132","414_BS911-50","422_BS911-58")
cols_to_drop <- c("SY_Tepee","1-6","10.1618","Buck","DH130935")

wmb_gt <-wmb_gt[,!(names(wmb_gt) %in% cols_to_drop)];wmb_gt<-wmb_gt %>% dplyr::rename( FLAVIA=AC_Flavia,SCALA="KWS_Scala",TEPEE="SY-Tepee",ENDEAVOR=Endeavor,WINTMALT=Wintmalt)
#colnames(wmb_gt)[c(12:672)]<-sapply(strsplit(colnames(wmb_gt)[12:672],"_"), "[", 2)
# we have some lines that needed to re genotyped, need to filter old ones out.
duplicate_names<-colnames(wmb_gt[which(duplicated(colnames(wmb_gt)))]);duplicate_names

AlaT_markers_1<-AlaT_markers_1[colnames(AlaT_markers_1)%in%colnames(wmb_gt)]

wmb_gt<-wmb_gt%>%add_row(AlaT_markers_1)%>%arrange(chrom,pos)

#add morex positions

#should be the correct alignment
Morex2019<-read.delim("data/Genotype_data/barley50kMarkerPositions_Morex2019Assembly.gff",header = FALSE)
#%>%as.data.frame()%>%dplyr::select(1,2,4,7,9)%>%
Morex2019<-Morex2019%>%as.data.frame()%>%dplyr::select(1,2,4,7,9)%>%dplyr::rename(Chrom=1,Type=2,pos=3,strand=4,Marker=5)
Morex2019$Chrom<-str_split_fixed(Morex2019$Chrom,"r" , 2)[,2]
Morex2019$Marker<-str_split_fixed(Morex2019$Marker,"=" , 2)[,2]
Morex2019=Morex2019%>%arrange(Chrom,pos)

#check for NAs
table(Morex2019$Chrom,useNA = "ifany")
myGM20=data.frame(SNP=wmb_gt$`rs#`)
myGM20[myGM20$SNP=="Qsd1",]
myGM20$Chromosome = Morex2019$Chrom[match(myGM20$SNP, Morex2019$Marker)]; myGM20$Chromosome = substr(myGM20$Chromosome ,1,1)
myGM20$Chromosome[which(myGM20$Chromosome=="U")]="8"
myGM20$Position = as.character(Morex2019$pos[match(myGM20$SNP, Morex2019$Marker)])
#add Qsd1 position
myGM20[myGM20$SNP=="Qsd1",]
myGM20[9002,]<-c("Qsd1","5",442160000)
myGM20<-myGM20[with(myGM20, order(Chromosome,Position)),]

library(GAPIT3)

wmb_gt$chrom<-as.numeric(sapply(str_split(wmb_gt$chrom,"H"),"[",1))
wmb_gt[wmb_gt$chrom==0,]$chrom<-8
wmb_gt$chrom<-as.integer(wmb_gt$chrom)
wmb_gt$pos<-as.integer(wmb_gt$pos)
wmb_gt=wmb_gt%>%arrange(chrom,pos)

#myG <- read.table("~/Downloads/GAPIT_Tutorial_Data/mdp_genotype_test.hmp.txt", head = TRUE)
##myGAPIT <- GAPIT(G=myG, output.numerical=TRUE)
wmb_gt_t<-wmb_gt

wmb_gt_t$chrom = myGM20$Chrom[match(wmb_gt_t$`rs#`, myGM20$SNP)];wmb_gt_t$chrom<-as.integer(wmb_gt_t$chrom)

wmb_gt_t$pos = myGM20$Position[match(wmb_gt_t$`rs#`, myGM20$SNP)];wmb_gt_t$pos<-as.numeric(wmb_gt_t$pos)            
wmb_gt_t$pos 
wmb_gt_t<-wmb_gt_t%>%tidyr::drop_na(chrom)%>%dplyr::arrange(chrom,pos)

wmb_gt_t=wmb_gt_t%>%as.data.frame()%>%add_row(wmb_gt_t[1,],.before = 1)
wmb_gt_t[1,]$"rs#"<-"JHI-Hv50k-2016-6205"
wmb_gt_t[1,]$pos<-3470194-1
library(GAPIT3)
GAPIT_wmb<- GAPIT(G=wmb_gt_t%>%as.data.frame(), output.numerical=TRUE,file.output = FALSE)

names(GAPIT_wmb$GD) <- gsub(x = names(GAPIT_wmb$GD), pattern = "\\.", replacement = "-")

library(rrBLUP)
wmb_GD<-GAPIT_wmb$GD[-1];wmb_GD=as.data.frame(t(wmb_GD))
wmb_GD=as.data.frame(apply(wmb_GD,1, function(x) { 
  x[which(x == 0)] = -1
  x[which(x == 1)] <- 0
  x[which(x == 2)] = 1
  
  
  
  x
}))
library(rrBLUP)
wmb_GD=rrBLUP::A.mat(wmb_GD, return.imputed = T,impute.method = "EM")

names(wmb_GD) <- gsub(x = names(wmb_GD), pattern = "\\.", replacement = "-") 
wmb_GD_rr<-wmb_GD



Gi<-wmb_GD_rr$imputed

rownames(wmb_GD_rr$A)<-toupper(rownames(wmb_GD_rr$A))
colnames(wmb_GD_rr$A)<-toupper(colnames(wmb_GD_rr$A))
library(pedigree)
#were only using pedigree info here
load("data/Phenotype_Data/Field_data_2020_2022.Rdata")

ped<-all_pheno%>%dplyr::select(GID,family,Check)%>%unique()
ped[!ped$Check==0,]$family<-toupper(ped[!ped$Check==0,]$family)
ped$parent1<-sapply(str_split(ped$family,"/"),`[`,1)
ped$parent2<-sapply(str_split(ped$family,"/"),`[`,2)
ped[!ped$Check==0,]$parent1<-"FOUNDER"
ped[!ped$Check==0,]$parent2<-"FOUNDER"
ped[ped$Check==3,]$parent1<-"CHECK"
ped[ped$Check==3,]$parent2<-"CHECK"
ped$parent1<-toupper(ped$parent1)
ped$parent2<-toupper(ped$parent2)
ped[grep("SY",ped$parent1),]$parent1<-"TEPEE"
ped[grep("SY",ped$parent2),]$parent2
ped[!ped$Check==0,]$parent1<-0
ped[!ped$Check==0,]$parent2<-0

#ped_ng<-ped%>%filter(!ped$GID%in%colnames(Am)|!ped$Check==0)

wmb_ped<-ped%>%dplyr::select(-c(2,3))%>%unique()%>%dplyr::rename(Ind=1,Par1=2,Par2=3)
wmb_ped$Population<-"Recombinant_inbred_lines"
wmb_ped[grep("BS",wmb_ped$Ind),]$Population<-"Double_haploid"
wmb_ped[grep(0,wmb_ped$Par1),]$Population<-"Parents"
wmb_ped[grep("END",wmb_ped$Ind),]$Population<-"Check"
wmb_ped[grep("CHA",wmb_ped$Ind),]$Population<-"Check"
#Should do LD pruning here
library(gdsfmt)
library(SNPRelate)







save(wmb_GD_rr,file="data/Genotype_data/wmb_GD_rrblup.Rdata")

save(GAPIT_wmb,file = "data/Genotype_data/GAPIT_wmb.Rdata")
save(wmb_ped,file="data/Genotype_data/wmb_pedigree.Rdata")




#snpgdsClose(gdsfile)


pca <- snpgdsPCA(gdsfile)
pca$varprop
pc.percent <- pca$varprop*100
head(round(pc.percent, 2))
plot(1:10,pc.percent[1:10])

rownames(wmb_GD)

WinterRelationship = wmb_GD_rr$A
#list_in<-wmb_ped%>%filter(!Population%in%c("Recombinant_inbred_lines"))# we are just doing the non RIL lines

#WinterRelationship=WinterRelationship [rownames(WinterRelationship)%in%list_in$Ind,colnames(WinterRelationship)%in%list_in$Ind]
WinterPCA = eigen(WinterRelationship)
WinterPVEPCA = WinterPCA$values/sum(WinterPCA$values)
data.frame(ordinal = 1:10, PVE = WinterPVEPCA[1:10]) %>%plot(., xlab = 'PC', col = 'red') 

wmb_GD$taxa<-rownames(wmb_GD)

t<-wmb_GD$imputed%>%as.data.frame()%>%tibble::add_column(taxa=rownames(.),before=1)

#18-Scala x Tepee, 19 Scala x Wintmalt, 20 Scala x Flavia, 21 Scala x DH130910, 22, Tepee x Wintmalt
#23 Flavia x SY Tepee 24 Tepee x DH130910, 25 Flavia x Wintmalt, 26 Wintmalt x DH130910, 27 Flavia x DH130910

#t<-t%>%mutate(Family = plyr::mapvalues(family, from = c("18","19","20","21","22","23","24","25","26","27",'BS6','BS7','BS8','BS9','DH1','Fla','Tep','Sca','Win'), 
 #                   to = c("SY_Tepee/Scala","Scala/Wintmalt","Scala/Flavia","Scala/DH130910","SY_Tepee/Wintmalt","SY_Tepee/Flavia","SY_Tepee/DH130910","Flavia/Wintmalt","Wintmalt/DH130910","Flavia/DH130910",'Flavia/DH130910','Scala/DH130910','SY_Tepee/DH130910','Wintmalt/DH130910',
 #                        'DH130910','Flavia/DH130910','SY_Tepee/DH130910','Scala/DH130910','Wintmalt/DH130910')))
t$GID<-rownames(t)
t$GID
table(wmb_ped$Population)


t<-t%>%mutate(Family=plyr::mapvalues(rownames(.),from = ped$GID,to=ped$family))
t$Family
winterlinePCAvalues = WinterPCA$vectors %>% data.frame()%>% 
  mutate(family =t$Family,
         taxa = rownames(t),
         shapes = ifelse(taxa %in% c('DH130910', 'FLAVIA','TEPEE','WINTMALT','SCALA'), taxa, 'Lines'),
         size = ifelse(taxa %in% c('DH130910', 'FLAVIA','TEPEE','WINTMALT','SCALA','ENDEAVOR'), 4, 3))
winterlinePCAvalues
library(ggplot2)
winterlinePCAvalues %>% ggplot(aes(x = X1, y = X2, color = family)) + geom_point()+
  winterlinePCAvalues%>% ggplot(aes(x = X1, y = X3,color = family)) + geom_point()+
  winterlinePCAvalues%>% ggplot(aes(x = X2, y = X3,color = family)) + geom_point()

winterlinePCAvalues  %>%
  ggplot(aes(x = X1, y = X2, color = family, shape = shapes)) + geom_point(aes(size = size))+
  ggtitle(label ="Principal Component Analysis \n of the WMB Population")+
  xlab('PC1')+ylab('PC2')+
  
  theme_bw()


dev.off()
#myGenoFile <- system.file("extdata", "mdp_genotype_test.hmp.txt.gz", package = "GAPIT3")
#myGenotypes  <- read.table(myGenoFile, header = FALSE)

```